Bayesian spectral density estimation using P-splines with quantile-based knot placement

### 2020-11-04 responses/responses3/response.tex

	\response{This could be a potential avenue for future research, however, we think it is beyond the scope of this paper.  An analysis of those characteristics can be quite extensive and could constitute a paper in itself, see for instance \cite{Likhachev2017}.  We would like to highlight that the number of knots is considered non-important by many, since the penalty parameter plus the penalty order difference penalties specified in the penalty matrix $\textbf{P}$ penalise the inclusion of B-spline densities, controlling the smoothness of the estimate.  This means that we only need to select a minimum value, the penalty terms take care of the rest. To quote \cite{Eilers2015}: ``With the number of B-splines in the basis we can tune the smoothness of a curve to
the data at hand. A smaller number of splines gives a smoother result. However, this is not the only possibility. We can also use a large basis and additionally constrain the
coefficients of the B-splines, to achieve as much smoothness as desired. A properly
chosen penalty achieves this"; and also ``The number of B-splines can be (much) larger than the number of observations. The
penalty makes the fitting procedure well-conditioned. This should be taken literally:
even a thousand splines will fit ten observations without problems. Such is the power
of the penalty."}\bigskip

	Other minor comments: \bigskip

	\item \refereeQuote{Page 2, line 17, what is $\lambda_i$?}\smallskip

	\response{$\lambda_i$ denote the Fourier frequencies. These have now been defined just after the first equation on page 2.} \bigskip

@@

	\item \refereeQuote{ Page 4, lines 5-7, ``In particular, the approximation error of the mixture of B-spline distributions can be made arbitrarily small by finding suitable knot locations and increasing the number of knots", what does ``made arbitrarily small" mean? Do you mean the approach with the beta densities the error cannot be made arbitrarily small?}\smallskip

	\response{Arbitrarily small means that the approximation error converges to zero with increasing number of knots. We have  cited the paper by \cite{Perron:2001} that make this statement precise. But we do not mean to imply that this is not the case for the mixture of beta densities. Of course, according to the well-known Weierstrass approximation theorem, any continuous function can be approximated in the supnorm by a linear combination of Bernstein polynomials. If the continuous function is a density, the approximant becomes a mixture of beta distributions. By increasing the number of beta distributions, the approximation error can be made arbitrarily small. But as shown in \cite{Perron:2001} the rate of approximation is faster for Bsplines than for Bernstein polynomials. We have changed this paragraph in the manuscript accordingly.}

@@

	\response{Fixed.} \bigskip

	\item \refereeQuote{Page 5, line 20, ``If $\tau = \int_0^1f(\tau \omega)d \omega$ is the normalizing constant", it implies $s_r$ is a density, is it correct?}

	\response{Yes, $s_r$ is a pdf.  Note that it is actually a mixture of B-spline densities.}
\end{enumerate}

\subsection*{Further Changes}
Here some extra changes:
\begin{enumerate}
	\item We have defined $K^*$ as the number of internal knots (B-splines and BspDP prior section).
	\item We have specified the prior on $\tau$ (P-spline prior subsection of section 2). 
	\item We have added $s_r(\omega)$ in Eq(3), short notation used in the posterior distribution of~$\tau$ (Posterior computation section).	
	\item We have slightly simplified our knot location approach.  This has led to minor variations in the results shown in the application section, which has been updated.  This change has not impacted the conclusions.  This version of the knot location algorithm is described step by step in the Quantile-based knot placement section.
\end{enumerate}

### 2020-11-04 pSplines.tex
update

As the periodogram fluctuates around the true spectral density, most  nonparametric methods for spectral density estimation hinge on  smoothing the periodogram
\begin{align*}
I_n(\lambda_l) = \dfrac{1}{2 \pi n} \left| \sum_{t=1}^{n} Y_t \exp \left( -i t \lambda_l\right)\right|^2,
\end{align*}
in one way or another, where $ \lambda_l=2\pi/n$ for $l=0,\ldots,\nu = \left\lfloor (n-1)/2 \right\rfloor$ are the Fourier frequencies and $\bm{Y}=(Y_1,\ldots,Y_n)$ is a stationary time series of length $n$. {\pmr The periodogram ordinates evaluated at the Fourier frequencies  are asymptotically independent and follow exponential distributions with mean $f(\lambda_l)$.  Thus, they are asymptotically  unbiased estimates of $f(\lambda_l)$  but not consistent \citep{Brockwell:1986}.}

@@

Our knot location proposal works as follow.  We calculate the periodogram and apply a square root transformation in order to have more regular magnitudes and eliminate a potential trend.  Second, we standardize these values, apply the absolute value function,{\pmr and normalise them}.  
We treat this transformed periodogram as a {\pmr probability mass} function (pmf) and interpolate its distribution function \textcolor{red}{ by the continuous cumulative distribution function} $\widetilde{F}$.
Finally, the knots are allocated according to quantiles of $\widetilde{F}$.
 Thus, in those areas in which the periodogram has sharp and abrupt changes, our procedure will assign more knots in proportion to their magnitudes. Hereafter, we will refer to this scheme as {\em  quantile-spaced (Q-spaced) knots}.%\textit{periodogram-spaced knots} or short \textit{P-spaced knots}.
%REMARK: HOW ABOUT "PERIODOGRAM-SPACED KNOTS"	OR 'P-SPACED KNOTS'  INSTEAD OF DISTRIBUTED KNOTS?

@@

{\pmr
The procedure for the frequencies in the unit interval $\omega_l = 2 l/n  \in[0,1]$, which can be easily translated to the frequency domain, can be summarized as follows:	
\begin{enumerate}
	\item Let $\bm{x} = \left (\sqrt{I_n(\pi\omega_1)}, \dots, \sqrt{I_n(\pi\omega_{\nu})} \right)$ be the square root of the periodogram values.
	\item Define $y_l = \left|\dfrac{x_l - \bar{x}}{S_x} \right|$ where $\bar{x}$ denotes the average and $S_x$ the standard deviation of ${\bm x}$.
	\item Define $z_l = \frac{y_l}{\sum_{l=1}^{\nu} y_l}$, $Z_l= \sum_{j\leq l} z_j$, and let $\widetilde{F}$ denote the continuous cdf that interpolates the $Z_l$, $l=1,\ldots,\nu$.
	\item Let $\bm{q} =\left( \frac{1}{K^*+1}, \frac{2}{K^*+1}\dots, \frac{K^*}{K^*+1} \right)$ be a vector of $K^*$ equally spaced points  in  $[0,1]$.
	\item The Q-spaced internal knot vector is given by the corresponding quantiles $\bm{\xi}^* = \left(\widetilde{F}^{-1}(q_1),\dots, \widetilde{F}^{-1}(q_{K^*}) \right)$. %where the inverse function is obtained by interpolation.	
\end{enumerate}

>>>>

As the periodogram fluctuates around the true spectral density, most  nonparametric methods for spectral density estimation hinge on  smoothing the periodogram
\begin{align*}
I_n(\lambda_l) = \dfrac{1}{2 \pi n} \left| \sum_{t=1}^{n} Y_t \exp \left( -i t \lambda_l\right)\right|^2,
\end{align*}
in one way or another,{\pmr where $ \lambda_l=2\pi l/n$ for $l=0,\ldots,\nu = \left\lfloor (n-1)/2 \right\rfloor$ are the Fourier frequencies} and $\bm{Y}=(Y_1,\ldots,Y_n)$ is a stationary time series of length $n$. The periodogram ordinates evaluated at the Fourier frequencies  are asymptotically independent and follow exponential distributions with mean{\pmr$f(\lambda_l)$}.  Thus, they are asymptotically  unbiased estimates of{\pmr$f(\lambda_l)$}  but not consistent \citep{Brockwell:1986}.

@@

Our knot location proposal works as follow.  We calculate the periodogram and apply a square root transformation in order to have more regular magnitudes and eliminate a potential trend.  Second, we standardize these values, apply the absolute value function,{\pmr and normalise them}.  
We treat this transformed periodogram as a{\pmr probability mass} function (pmf) and interpolate its distribution function \textcolor{red}{ by the continuous cumulative distribution function} $F$.
Finally, the knots are allocated according to the quantiles of $F$.
 Thus, in those areas in which the periodogram has sharp and abrupt changes, our procedure will assign more knots in proportion to their magnitudes. Hereafter, we will refer to this scheme as {\em  quantile-spaced (Q-spaced) knots}.%\textit{periodogram-spaced knots} or short \textit{P-spaced knots}.
%REMARK: HOW ABOUT "PERIODOGRAM-SPACED KNOTS"	OR 'P-SPACED KNOTS'  INSTEAD OF DISTRIBUTED KNOTS?

@@

{\pmr
The procedure for the frequencies in the unit interval $\omega_l = 2 l/n  \in[0,1]$, which can be easily translated to the frequency domain, can be summarized as follows:	
\begin{enumerate}
	\item Let $\bm{x} = \left (\sqrt{I_n(\pi\omega_1)}, \dots, \sqrt{I_n(\pi\omega_{\nu})} \right)$ be the square root of the periodogram values.
	\item Define $y_l = \left|\dfrac{x_l - \bar{x}}{S_x} \right|$ where $\bar{x}$ denotes the average and $S_x$ the standard deviation of ${\bm x}$.
	\item Define $z_l = \frac{y_l}{\sum_{l=1}^{\nu} y_l}$, $Z_l= \sum_{j\leq l} z_j$, and let $F$ denote the continuous cdf that interpolates the $Z_l$, $l=1,\ldots,\nu$.
	\item Let $\bm{q} =\left( q_1, \dots, q_{K^*} \right)$ be a vector of $K^*$ equally spaced points  in  $[0,1]$, that is $q_j = \frac{j-1}{K^*-1}$, for $j=1,\dots,K^*$.
	\item The Q-spaced internal knot vector is given by the corresponding quantiles $\bm{\xi}^* = \left(F^{-1}(q_1),\dots, F^{-1}(q_{K^*}) \right)$. %where the inverse function is obtained by interpolation.	
\end{enumerate}
